Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism

### 2025-08-25 Manuscript/Ant_Cactus_RESPONSE.tex
Updates
Manuscript Response

\comment{Like the reviewers, especially Reviewer \#1, I could also have used a little more detail on some of the data collection methods. Reviewer \#1 notes that little information is provided on how seed production was estimated, and I agree. For my part, I could also have used a little more detail on how floral viability was measured. The methods say “we distinguished between flowerbuds that were viable versus aborted” but I wasn’t clear how.}
\response{Thank you for these suggestions. 
	We agree that methods for estimating seed production were a little under-developed in our previous submission because we relied too heavily on citations of our prior work. 
	We now describe these methods more thoroughly (lines \ref{}). 
	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{}). 
	We have gone through the data collection methods section and added greater detail throughout (lines \ref{}), which we hope will make our work more understandable and reproducible.}

\comment{I also encourage the authors to pay close attention to responding to several important points raised by Reviewer \#2, especially with regards to 1) what evidence for the portfolio effect would look like in a mutualism in which partners participate sequentially, and 2) providing the reader with a little more ‘hand holding’ when it comes to the rather complex model. Like the reviewer, I cannot claim that I would spot an error in the model if there were one, and it may be helpful to build up to the full Integral Projection Model in more stages, as Reviewer \#2 very helpfully suggests. }
\response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines XX and Figure XX). 
	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{}). 
	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{}). }

@@

\comment{There is very little mention of some of the supplemental results in the main text (e.g., the herbivory analysis). Please make sure all supplemental results are clearly described in the Results section of the main text.}
\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{}. \ref{}).}

>>>>

\comment{Like the reviewers, especially Reviewer \#1, I could also have used a little more detail on some of the data collection methods. Reviewer \#1 notes that little information is provided on how seed production was estimated, and I agree. For my part, I could also have used a little more detail on how floral viability was measured. The methods say “we distinguished between flowerbuds that were viable versus aborted” but I wasn’t clear how.}
\response{Thank you for these suggestions. 
	We agree that methods for estimating seed production were a little under-developed in our previous submission because we relied too heavily on citations of our prior work. 
	We now describe these methods more thoroughly (lines \ref{ms-res:seed_prod}). 
	We distinguished between viable and aborted flowerbuds through visual assessment, also clarified in our revision (line \ref{ms-res:viable}). 
	We have gone through the data collection methods section and added greater detail throughout (lines \ref{ms-res:repro1}, \ref{ms-res:repro2},  \ref{ms-res:viable}, \ref{ms-res:exhaustive}), which we hope will make our work more understandable and reproducible.}

\comment{I also encourage the authors to pay close attention to responding to several important points raised by Reviewer \#2, especially with regards to 1) what evidence for the portfolio effect would look like in a mutualism in which partners participate sequentially, and 2) providing the reader with a little more ‘hand holding’ when it comes to the rather complex model. Like the reviewer, I cannot claim that I would spot an error in the model if there were one, and it may be helpful to build up to the full Integral Projection Model in more stages, as Reviewer \#2 very helpfully suggests. }
\response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines \ref{ms-res:portfolio_evidence} and Figure \ref{ms-app:portfolio}). 
	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{ms-res:portfolio_asynchronous}). 
	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{ms-res:portfolio_methods}). }

@@

\comment{There is very little mention of some of the supplemental results in the main text (e.g., the herbivory analysis). Please make sure all supplemental results are clearly described in the Results section of the main text.}
\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{res:portfolio_methods}, \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}

### 2025-08-25 Manuscript/ant_cactus_SUBMISSION2.tex
Updates
Manuscript Response

\newpage{}
\section*{Introduction} \label{res:introduction}
%% introduce multispecies mutualisms and sequential mutualisms with definitions
\revise{Mutualisms, species interactions where all participants receive net benefits, are key components of ecological communities and can enhance the fitness and population growth rates of the partners involved.  \linelabel{res:27}
While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal species interacts with multiple partner species.
Growing evidence suggests that pairwise interactions may be poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}. 
A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species). 
Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}. 
Partner diversity can arise either from simultaneous interactions or sequential ones, where individual hosts experience turnover in partner identity over time \citep{Djieto-Lordon2005, Ness2006}. \linelabel{res:L22}
Understanding how partner diversity and turnover affect the overall fitness benefits of mutualism is an open question -— especially in the context of environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}. \linelabel{res:L39}}

@@

\subsection*{Data Collection}
This study is based on long-term data spanning 2004 to 2023 that documented plant demographic rates in relation to their ant partner status. 
From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown). 
For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. 
We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (aborted flowerbuds are distinguishable by dry and withered true leaves and petals). 
We recorded the ant species presence and identity (or vacancy if no ants present).
Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species. 
\linelabel{res:exhaustive} Each year, the field team searched exhaustively for new recruits by walking slowly in a standard formation across the plot area, and these were added to the census (cactus seedlings are small ($<2 cm$ in height) and cryptic so it is likely that some are missed even with a standardized search protocol). 
These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
\tom{In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. }{Update with new data package citation, and also update data availability statement.}
In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds (and therefore are missing some vital rate information for those years).

@@

\subsubsection*{Mathematical formulation}
Mathematically, we modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates. 
The model also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
Seed bank dynamics are given by:
\begin{linenomath*}
	\begin{gather}
B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
B^2_{t+1} =  (1 - \gamma_1)B^1_{t}
	\label{eqn:IPM1}
	\end{gather}
\end{linenomath*}
\noindent %In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$. 
The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability, and plants in different ant states can fluctuate independently.
% (good years for \textit{L. apiculatum} -occupied plants may be bad years for \textit{C. opuntiae}-occupied plants, for example). 
Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
Seeds are multiplied by the probability of escaping post-dispersal seed predation ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.

@@

\begin{figure}[H]
\includegraphics[width=0.75\linewidth]{Figures/survival_plot.png}
\caption{Probability of survival in relation to size and ant partner. Layout as in Fig. \ref{fig:Grow}, except here points show data as binned means, where point size is proportional to number of observations. Grey areas show the 90\% confidence interval for the mean.}
\label{fig:Surv}
\end{figure}

Ant visitation was associated with increased floral viability and ant identity determined the strength of viability benefits.
Mean viability rates spanned 55--81\% (Figure \ref{fig:Viab}).
\textit{L. apiculatum}-tended plants had the highest mean viability rate (86.1\% [95\% credible interval: 77.6--92.4\%]), while there were similar viability rates for vacant (60.0\% [44.3--75.0\%]), Other-tended (60.6\% [43.7--75.5\%]), and \textit{C. opuntiae}-tended plants (57.1\% [40.6--72\%]). 
\tom{Furthermore, \textit{C. opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L. apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}.}{I think a reviewer asked to see this result, so be sure to reference the figure in the appendix and flag this as a revision in the response letter. Also, the figure legend should be updated to indicate that we are looking at posterior probability distributions, which some readers will not understand if you don't explicitly state.}

@@

Partner identity and composition were not strongly consequential for plant fitness (Figure \ref{fig:LambdaMeans} b-c). 
Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans} b). 
While a hypothetical \textit{L. apiculatum}-only population had the highest mean fitness, it was not significantly higher than \textit{C. opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) only populations. 
Furthermore, the fitness of \textit{L. apiculatum}-only was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$). 
The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above.
%Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
%Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.

@@

% Value of Long-term data
This study highlights the value of long-term data in investigating species interactions. 
Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \citep{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
Separate studies have analyzed how inter-annual variability impacts focal mutualists \citep{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
The long term data set we used gave us the ability to consider, for the first time, the combined effects of partner identity, partner turnover, and temporal stochasticity. 
By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of ant visitation. 
For example, our previous study that focused only on reproductive vital rates suggested that \textit{C. opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed \tom{set}{Reference the appendix figure here.} \citep{Ohm2014}. 
Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.

% Limitations
As with any study, there are limitations to consider when interpreting our results.
First, we used observational data to infer ant effects on plant demography. 
However, we have previously conducted experimental manipulations which revealed that ant presence has causal effects on plant demographic rates through anti-herbivore defense \citep{Miller2007,Ohm2014}.
The combination of long-term observational data backed up by experimental results gives us greater confidence in our causal interpretations. 
Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
\tom{Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction. }{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.}
Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences. 
As more long-term data accumulate, we may gain more information with which to further unpack this group of species.

@@

>>>>

\newpage{}
\section*{Introduction} \label{res:introduction}
%% introduce multispecies mutualisms and sequential mutualisms with definitions
\revise{Mutualisms, species interactions where all participants receive net benefits, are key components of ecological communities and can enhance the fitness and population growth rates of the partners involved.  \linelabel{res:27}
While traditionally framed as pairwise (+/+) interactions, mutualisms are often embedded within multispecies communities in which a focal species interacts with multiple partner species.
Growing evidence suggests that pairwise interactions may be poor predictors of the net effects of multi-species mutualism \citep{Afkhami2014,Palmer2010,Bascompte2009,Dattilo2014}. 
A focal mutualist may interact with multiple guilds of partner types (e.g., plants that interact with pollinators, seed dispersers, soil microbes, and ant defenders) or with multiple partner species within the same guild (e.g., plants visited by multiple pollinator species). 
Within a mutualist guild, partner species often differ in the amount or type of goods or services they provide, making partner identity an important source of contingency in mutualism \citep{Stanton2003}. 
Partner diversity can arise either from simultaneous interactions or sequential ones, where individual hosts experience turnover in partner identity over time \citep{Djieto-Lordon2005, Ness2006}. \linelabel{res:L22}
Understanding how partner diversity and turnover affect the overall fitness benefits of mutualism is an open question, especially in the context of environmental variability, where different partners may confer benefits unevenly across space or time  \citep{rogers2014, Thibaut2012, Frederickson2005, Palmer2010}. \linelabel{res:L39}}

@@

\subsection*{Data Collection}
This study is based on long-term data spanning 2004 to 2023 that documented plant demographic rates in relation to their ant partner status. 
From 2004 to 2008, we censused 134 plants distributed across three spatial blocks. 
This initial census group was discontinued in 2009, when we established six 30 $\times$ 30-meter plots and tagged all tree cholla within those plots. 
Two additional 30 $\times$ 30-meter plots were added in 2011, and this group of eight plots (mean $\pm$ SD Tree cholla per plot: 87.30 $\pm$ 18.28) was censused annually through 2023 (with the exception of 2020 due to the pandemic shutdown). \linelabel{res:repro1}
For all plants, in May or early June of each year we recorded plant survival since the last survey (if the plant had any green flesh it was considered alive) and, for survivors, we recorded height (cm), maximum crown width (cm), and crown width perpendicular to the maximum (cm) (sections with green were measured for size).
Size measurements were used to calculate plant volume ($cm^3$) based on the volume of an elliptical cone. \linelabel{res:repro2}
We measured reproduction by counting flowerbuds, and in most years we distinguished between flowerbuds that were viable versus aborted (aborted flowerbuds are distinguishable by dry and withered true leaves and petals).  \linelabel{res:viable}
We recorded the ant species presence and identity (or vacancy if no ants present).
Occurrences of more than one ant species on one plant were rare (less than 5\% of observations), and for the purpose of this analysis we classified the plant as being occupied by the more abundant species. 
\linelabel{res:exhaustive} Each year, the field team searched exhaustively for new recruits by walking slowly in a standard formation across the plot area, and these were added to the census (cactus seedlings are small ($<2 cm$ in height) and cryptic so it is likely that some are missed even with a standardized search protocol). 
These data allowed us to link each plant's demographic fate (survival, growth, and reproduction) to its state of ant visitation. 
\tom{In total, the data set includes a total of 9,787 observations of 1,141 individuals across 15 complete transition years (spanning May/June of year $t$ to May/June of year $t+1$) \citep{DataCholla}. }{Update with new data package citation, and also update data availability statement.}
In addition to missing the year 2020, there are gaps in the time series where we switched plots or plants (and thus broke up transition years for growth/survival and partner turnover) or where we did not distinguish between viable and aborted flower buds (and therefore are missing some vital rate information for those years).

@@

\subsubsection*{Mathematical formulation}
Mathematically, we modeled the tree cholla life cycle using continuously size-structured plants where number of plants of size $x$  and ant state $a$ in year $t$ ($n(x,a)_{t}$) predicts the number of plants of size $x'$ and ant state $a'$ in year $t+1$ ($n(x',a')_{t+1}$) based on a size- and ant-specific vital rates. 
The model also includes two discrete seed banks ($B^1_{t}$ and $B^2_{t}$) corresponding to 1 and 2-year old seeds. 
Seed bank dynamics are given by:
\begin{linenomath*}
	\begin{gather}
B^1_{t+1} = \delta \sum_{a=1}^{A} \int_L^U  \kappa(a) P(x;\pmb{\tau}^P) F(x;\pmb{\tau}^F) V(a;\pmb{\tau}^V_{a}) n(x,a)_{t} dx \\
B^2_{t+1} =  (1 - \gamma_1)B^1_{t}
	\label{eqn:IPM1}
	\end{gather}
\end{linenomath*}
\noindent %In these equations $x$ and $x'$ indicate the size of a plant in years $t$ and $t+1$ respectively, $a$ and $a'$ indicate the ant partner of a plant in years $t$ and $t+1$.
Functions $P(x;\pmb{\tau}^P)$ and $F(x;\pmb{\tau}^F)$ give the probability of flowering in year $t$ and the number of flowerbuds produced in year $t$, respectively, by plants of size $x$ in year $t$. 
The proportion of flowerbuds that remain viable through fruit set ($V(a;\pmb{\tau}^V_{a})$) and the number of seeds per fruit ($\kappa(a)$) are dependent on ant state $a$. 
The vectors $\pmb{\tau}$ give year-specific deviates (mean zero) and appear in functions for which we can estimate temporal stochasticity from the long-term data; superscripts indicate the corresponding vital rate and, when present, the $a$ subscript indicates that deviates are specific to plants in ant state $a$.
For example, temporal deviates $\pmb{\tau}^V_{a}$ describe better- and worse-than-average years for flowerbud viability, and plants in different ant states can fluctuate independently.
% (good years for \textit{L. apiculatum} -occupied plants may be bad years for \textit{C. opuntiae}-occupied plants, for example). 
Seed production is integrated over the size distribution, from the lower $L$ to upper $U$ size limits, and summed over all possible ant states ($A=4$) giving total seed production. 
Seeds are multiplied by the probability of escaping post-dispersal seed predation ($\delta$) to give the number of seeds that enter the one-year old seed bank. 
Plants can recruit out of the year-one seed bank with probability $\gamma_1$ or transition to the two-year seed bank with a probability of $1 - \gamma_1$. 
Seeds in the two-year seed bank are assumed to either germinate with probability $\gamma_2$ or die.  \linelabel{res:seed_prod}

@@

\begin{figure}[H]
\includegraphics[width=0.75\linewidth]{Figures/survival_plot.png}
\caption{Probability of survival in relation to size and ant partner. These are the posterior probability distributions. Layout as in Fig. \ref{fig:Grow}, except here points show data as binned means, where point size is proportional to number of observations. Grey areas show the 90\% confidence interval for the mean.}
\label{fig:Surv}
\end{figure}

Ant visitation was associated with increased floral viability and ant identity determined the strength of viability benefits.
Mean viability rates spanned 55--81\% (Figure \ref{fig:Viab}).
\textit{L. apiculatum}-tended plants had the highest mean viability rate (86.1\% [95\% credible interval: 77.6--92.4\%]), while there were similar viability rates for vacant (60.0\% [44.3--75.0\%]), Other-tended (60.6\% [43.7--75.5\%]), and \textit{C. opuntiae}-tended plants (57.1\% [40.6--72\%]). 
Furthermore, \textit{C. opuntiae}-tended plants had fewer seeds per fruit (115.0[79.5--165.5] seeds) than vacant (147.2[114.1--189.9] seeds) or \textit{L. apiculatum}-tended plants (142.4[100.7--200.2] seeds), likely due to pollinator interference \citep{Ohm2014}. \linelabel{res:viability}
\ali{}{I am not sure where in the appendix you want this to be referenced?}

@@

Partner identity and composition were not strongly consequential for plant fitness (Figure \ref{fig:LambdaMeans} b-c). 
Among the one-partner scenarios, there was no strong evidence for any single best partner species (Figure \ref{fig:LambdaMeans} b). 
While a hypothetical \textit{L. apiculatum}-only population had the highest mean fitness, it was not significantly higher than \textit{C. opuntiae}-  ($Pr(\Delta\lambda^{L-C}_{S}=0.7)$) or Other- ($Pr(\Delta\lambda^{L-O}_{S}=0.69)$) only populations. 
Furthermore, the fitness of \textit{L. apiculatum}-only was consistent with all 2-partner scenarios ($Pr(\Delta\lambda_{S})$ ranged from 0.37 to 0.83).
However, as above, there was evidence for an opportunity cost of diversity wherein fitness of the three-partner scenario was lower than any of the 2-partner scenarios ($Pr(\Delta\lambda_{S})>0.95$). 
The lack of diversity benefits are not driven by the high overall frequency of \textit{L. apiculatum}. 
Using the simulations where all ants had equal frequencies across sizes (further explained in Appendix C), we found the same fitness patterns as described above. \linelabel{res:appendixc}
%Equal probability for transitioning into any ant state meant that the numbers of \textit{C. opuntiae} and Other ants were boosted significantly.
%Despite this, the fitness of scenarios including these less frequent ants were not increased meaningfully.

@@

% Value of Long-term data
This study highlights the value of long-term data in investigating species interactions. 
Previously studies have analyzed how partner identity and partner turnover impact focal mutualist fitness \citep{Fonseca2003, Dejean2008, Noe1994, Barrett2015, Bruna2014, Trojelsgaard2015}.
Separate studies have analyzed how inter-annual variability impacts focal mutualists \citep{Alonso1998, Alarcon2008, Ollerton2006, Horvitz1990, Lazaro2022}.
The long term data set we used gave us the ability to consider, for the first time, the combined effects of partner identity, partner turnover, and temporal stochasticity. 
By piecing together complete life cycle information from long-term data, we gain a more nuanced understanding of the fitness consequences of ant visitation. 
For example, our previous study that focused only on reproductive vital rates suggested that \textit{C. opuntiae} has overall parasitic fitness effects because activity of this species within tree cholla flowers can deter pollinators and reduce seed set (Figure \ref{app:seeds}) \citep{Ohm2014}. \linelabel{res:seeds}
Yet, the more complete analysis presented here, which accounts for reduced seed set alongside other demographic advantages in higher-sensitivity vital rates, indicates that this species is clearly a mutualist.

% Limitations
As with any study, there are limitations to consider when interpreting our results.
First, we used observational data to infer ant effects on plant demography. 
However, we have previously conducted experimental manipulations which revealed that ant presence has causal effects on plant demographic rates through anti-herbivore defense \citep{Miller2007,Ohm2014}.
The combination of long-term observational data backed up by experimental results gives us greater confidence in our causal interpretations. 
Second, we have not explicitly incorporated ant-herbivore interactions, even as these are the primary pathway through which ants influence plant demography. 
\tom{Surveys of herbivore damage from our long-term data are consistent with protective benefits of ant visitation (Figure \ref{app:herb}), and more explicit inclusion of plant-herbivore-defender interactions could be a valuable future direction. }{The reviews brought up that the herbivory figure in the appendix is not adequately discussed or contextualized, and I think that issue remains to be addressed.} \linelabel{res:herbivory}
Third, the rarity of several ant partner species imposed a practical constraint on assessing effects of partner diversity, since we needed to combine these into an ``Other'' category that might mask important differences. 
As more long-term data accumulate, we may gain more information with which to further unpack this group of species.

@@

\section*{Appendix E}\label{appendix:E}
To further evaluate the potential for a portfolio effect in this system, we conducted an additional analysis in which we artificially increased the degree of difference among ant partners. 
Specifically, we manipulated the correlation structure of ant-specific year random effects in the stochastic IPM. 
In the first scenario, we forced the correlation among ant species to be zero, representing complete independence of partner responses across years. 
In the second scenario, we forced the correlation as close to -1 as possible, representing maximally anti-correlated partner responses.

For each of these scenarios, we quantified the effect of partner diversity as the difference between projected population growth with and without partner diversity ($\lambda_S^4 - \lambda_S^0$), and compared results under synchronous versus non-synchronous environmental variation. 
Both scenarios resulted in greater benefits of partner diversity under Non-synchronous conditions, though the evidence was much stronger when the ants were anti-correlated.

This shows that portfolio effect is possible in this system if the ant partners diverged more greatly in their responses to annual variation. \linelabel{res:portfolio_evidence}

\begin{figure}
	\includegraphics[width=0.91\linewidth]{Figures/cor0_Portfolio_Effect.png}
	\caption{The mean estimations of $\lambda_{S}^4-\lambda_{S}^0$ for the non-synchronous and synchronous models are shown in green and pink respectively. }
	\label{app:portfolio}
\end{figure}
